### 2015-09-13 paper-1/main.tex
Yeah.

Consider the Phase II clinical trial that investigates the efficacy of 
iniparib in patients with breast cancer gene-associated (BRCA) ovarian 
cancer [\ref{Sanofi 2013}]. In the first stage 12 patients were treated with 
iniparib (BSI-201/SAR240550). The endpoint was defined by Response Evaluation 
Criteria in Solid Tumor (RECIST). If at least two Stage I patients succeeded,
then the trial would proceed to the next stage where additional patients were 
treated. If fewer than two of the 12 Stage I patients is successful then 
the trial would terminate.

Our goal is to know the distribution of the sample size of the design for 
planning purposes and to characterize the uncertainty in the estimate of the
success probability.  If all 12 patients are enrolled at once, 
as in the classic 
design, then the sample size is 12. However, in most clinical trials, the 
patients are enrolled sequentially, often with one patient's outcome realized 
before the next one enters the trial. In the present example, observing two 
successful patients allows us to proceed to the next stage. Similarly 11 
observed treatment failures also ends the stage. This sampling mechanism, in 
which the experiment ends as soon as any of the endpoints is reached, is 
call {\em curtailed sampling}. Under curtailed sampling the range of the 
sample size is between two and 12.

We will assume each of the patient enrollments can be modeled as independent, identically distributed Bernoulli($p$) random variables. The realization of the trial is a sequence of these random variables that stops when either a specified number of success or failures is reached. As an example, suppose in the previously described trial, two successes were reached after enrolling 10 patients (one in the third step and one at the $10^{th}$). The trial is visualized in Fig \ref{fig:kane_viz}. The vertical axis denotes the number of successful outcomes. The horizontal axis denotes the number of patients that have been enrolled. The horizontal and vertical line segments represent endpoints for the trial.

@@

Let $\,B_1, B_2, \ldots \,$ denote a sequence of independent, identically 
distributed, Bernoulli random variables with $\mathbb{P}\{B_i=1\}=p$, for 
probability parameter $0\leq p \leq 1$. In the clinical trial setting 
$\,B_i = 0$ corresponds to a successful enrollment. 
Let $s$ and $t$ be positive integers. 
Define the stopped negative binomial (SNB) random variable $Y$ as the smallest 
integer value such that $\,\{B_1, \ldots , B_Y\}\,$ contains {\em either} 
$\,s\,$ successes or else $\,t\,$ failures. The distribution of $\,Y\,$ has 
support on integer values in the range 
\begin{equation*}                                     %   (3)
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}
\end{equation*}
and it is distributed as
\begin{equation} \label{eqn:pmf}
\mathbb{P}\{Y=k\} = S(k, p, s)  \{s \leq k\} + T(k, p, t) \{t \leq k \} 
\end{equation}
where
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is the negative binomial probability of $s$ successes and
\begin{equation} \label{eqn:R}
T(k, p, t) = {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{equation}
is the negative binomial probability of $t$ failures. It should be noted that
this result expressed using the de Finetti where indicator functions and
sets are equivalent. A more complete explanation of this equivalence can be 
found in [\ref{Pollard 2002}].

@@

To illustrate that (\ref{eqn:pmf}) is the distribution of $Y$, consider the 
process $\mathbf{X} = \{X(k) : k \in \mathbb{Z}_{\ge 0} \}$ with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + B_{k+1} \left[ k-t < X_k < s \right]
\end{equation*}
where $B_{k+1}$ is once again distributed as Bernoulli($p$) for all $k$ in the domain of $\mathbf{X}$. The process can be conceptualized as a series of coin flips that stops when either $s$ heads or $t$ tails are reached. At each step a coin is flipped. If it is heads, the process advances one diagonally in the positive horizontal and vertical direction. Otherwise, it advances in the positive horizontal direction only. The process stops, $X_k$'s value does not change, after either $X_k = s$ or $X_k = k-t$. One example is given in Fig. \ref{fig:kane_viz}.

\begin{prop}
The distribution of the stopping time $\argminl_k [X_k \geq s, X_k \leq k-t]$ 
is equal to the SNB and is given in (\ref{eqn:pmf}). 
\end{prop}
\begin{proof}
The proof will proceed in two parts. First, a combinatorial justification will be given for the probability mass value on each element of the support. Second, it will be shown that the sum of the masses over the support sums to one.

The probability that a given realization of $\mathbf{X}$ reaches $s$ at 
time $k$ is  the probability that, at time $k-1$ there are $s-1$ successful 
coin flips and $k-s$ unsuccessful coin flips times the probability of a 
successful coin flip at time $k$ (\ref{eqn:N}). The probability $X_k = k-t$ 
is the probability that, at time $k-1$ there are $k-t$ successful coin flips 
and $t-1$ unsuccessful coin flips times the probability of an unsuccessful 
coin flip at time $k$ (\ref{eqn:R}).  Next, we need to show that the sum
\begin{equation} \label{eqn:sum_proof}
R = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{equation}
is equal to one.
First, substitute $i=k-s$ in the first summation and
$j=k-t$ in the second. 
\begin{equation*}
{j+s-1 \choose s-1} = {j+s-1 \choose j}
\end{equation*}
Then %(\ref{eqn:sum_proof}) becomes:
%\begin{equation*}
%R = \sum_{i=0}^{t-1} {i+s-1 \choose s-1} p^s (1-p)^i +
%\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
%\end{equation*}
$R$ can be written
as the cumulative distribution function of two 
negative binomial distributions
\begin{equation} \label{eqn:transformed_sum}
R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i \; + \;
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation}

Let $I_p(s, t)$ be the {\em regularized incomplete beta function}, which is 
also one minus the c.d.f. of the negative binomial distribution. This
function satisfies $I_p(s, t) = 1-I_{1-p}(t, s)$.  Then 
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-I_p(s, t) + 1 - I_{1-p}(t, s) \\
   &= 1. 
\end{align*}
\end{proof}
Proving that \ref{eqn:pmf} is a valid probability mass function.

From (\ref{eqn:pmf}) one can see that the mass at any point on the support 
of the distribution may come from either $X_k = s$, captured in the 
function $S$, or $X_k = k-t$, captured in the function $T$. The support of 
$S$ is defined by $s \leq S \leq s+t-1$ and the support of $T$ is defined 
by $t \leq T \leq s+t-1$.

@@

The probability mass function of $\,Y\,$ has a variety of shapes for different choices of the parameters $(s,\, t,\,p)$.
These shapes are illustrated in Fig.~\ref{shapes.fig}.
The SNB is related to the negative binomial distribution. Specifically, if 
$t$ is infinitely large then the $Y-s$ has a negative binomial distribution 
with
\begin{equation*}                                    %   (1)
   \mathbb{P}\{Y=s+j \}        \label{nb1.eq}
          = {{s+j-1}\choose{s-1}} p^s (1-p)^j
\end{equation*}
for $\,j=0, 1,\ldots\,$. A similar statement can be made when $s$ is large
and $t$ is small.

For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the 
riff-shuffle, or minimum negative binomial distribution~[\ref{Uppuluri 1970}].
Similar derivations of the closely-related maximum negative binomial discrete 
distributions also appear in~[\ref{Zhang 2000}, \ref{Zelterman 2004}]. However,
where the maximum negative binomial is the smallest number of coin-flips 
to observe at least $c$ successes {\em and} $c$ failures, the snb give
the number of coin flips to observe {\em either} $s$ heads or $t$ tails.

@@

\section{The Posterior Distribution}

@@

If we assume that $p$ has a Beta distribution, with constant prior parameters $\alpha$ and $\beta$ then we can calculate the closed form posterior 
distribution of the SNB as
\begin{align}
\mathbb{P} \{Y = k | \mathbf{B}, \alpha, \beta \} &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
\end{align}
where $B$ is the Beta function and $\mathbf{B}$ is the sequence of Bernoulli data.

When $p$ is known, the probability mass function (PMF) of the number of steps 
until the process is stopped is given by (\ref{eqn:pmf}). However, many times 
this is not the case. In Phase II clinical trials for example patients are 
often enrolled sequentially until there are enough patients to guarantee the 
desired level of power or until a specified number of adverse events are 
recorded. In cases such as these the endpoints $s$ and $t$ are known and we 
would like to be able to estimate $p$ either during enrollment to provide odds 
of successful enrollment or after an endpoint has been reached to determine 
the amount of uncertainty in the endpoint. 
%To do this, we will assume that $p$'s probability in the absence of 
%available data is $\text{Beta}(1/2, 1/2)$, the Jeffrey's prior. This 
%distribution was chosen for two reasons. First, it is the conjugate prior 
%of the (Negative) Binomial distribution and updating the distribution as data 
%are received is straightforward. Second, it is objective and limits the number 
%of ``prior coin-flips'' to one. This means that data are more heavily 
%influence the distributional characteristics of $p$ with more data providing 
%more proportional influence.

\begin{prop}
The posterior PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
\begin{align}
f(k | s, t, \alpha, \beta) &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
\end{align}
where $B$ is the Beta function.
\end{prop}
\begin{proof}
For notational simplicity, assume that $s,t \leq k \leq s+k-1$. When this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp
\end{align*}
The result follows by applying the definition of the Beta function to the integral terms.
\end{proof}

\section{A Bayesian Exploration Simon Two-stage Optimal Design}

Returning to the breast cancer trial introduced in the first section, consider
the first stage, having a sampel size of 12. A
response rate of 2/12 stage corresponds to an individual 
probability of responding of $p=0.1667$. The standard deviation of this 
probability is $\sqrt{p(1-p)/12}=0.1076$. If we were to use the Central 
Limit Theorem then the 95\% confidence interval around the mean would 
be $[-0.0442, 0.3776]$. While this approach is problematic since the number 
of samples (12 in this case) is small it does indicate how much uncertainty 
is built into the first stage. If we consider samples from both stages then 
the standard deviation drops to 0.0630 and the 95\% confidence interval 
would be $[0.0432, 0.2901]$.

These calculations expose two difficulties with the Simon Two-stage Optimal 
Design. First, success in the first stage is not a good indicator of success 
in the second stage and at the same time failure in first stage in no way 
implies failure in the second stage, especially when $p$ is close to it's 
threshold value. The variance in the estimate is so large that it is a poor 
estimator. The second difficulty is that the design is not {\em enrollment 
efficient}. That is, if the first stage is successful then additional patients 
are enrolled. Enrollment, especially for certain types of cancer, can be very 
difficult and can take time to find eligible patients. Furthermore, if 
all 12 of the patients in the first stage respond positively, should we 
really need to enroll additional patients?

@@

\item[\bibref{Pollard 2002}]
Pollard D. {\it A User's Guide to Measure Theoretic Probability}, 
Cambridge University Press, 2002.

>>>>

Consider the Phase II clinical trial that investigated the efficacy of 
iniparib in patients with breast cancer gene-associated (BRCA) ovarian 
cancer [\ref{Sanofi 2013}]. In the first stage 12 patients were treated with 
iniparib (BSI-201/SAR240550). The endpoint was defined by Response Evaluation 
Criteria in Solid Tumor (RECIST). If at least two Stage I patients succeeded,
then the trial would have proceed to the next stage where additional patients 
were treated. If fewer than two of the 12 Stage I patients is successful then 
the trial would have terminated.

Our goal is to describe the distribution of the sample size of the design for 
planning purposes and to use this distribution to estimate the
success probability.  If all 12 patients are enrolled at once, 
as in the classic 
design, then the sample size is 12. However, in most clinical trials, the 
patients are enrolled sequentially, often with one patient's outcome realized 
before the next one enters the trial. In the present example, observing two 
successful patients allows us to proceed to the next stage. Similarly 11 
observed treatment failures also ends the stage. This sampling mechanism, in 
which the experiment ends as soon as any of the endpoints is reached, is 
call {\em curtailed sampling}. Under curtailed sampling the range of the 
sample size is between two and 12.

We assume each of the patient outcomes can be modeled as independent, 
identically distributed Bernoulli($p$) random variables. The realization of 
the trial is a sequence of these random variables that stops when either a 
specified number of success or failures is reached. As an example, suppose in 
the previously described trial, two successes were reached after enrolling 10 
patients (one in the third step and one at the $10^{th}$). The trial is 
visualized in Fig \ref{fig:kane_viz}. The vertical axis denotes the number of 
successful outcomes. The horizontal axis denotes the number of patients that 
have been enrolled. The horizontal and vertical line segments represent 
endpoints for the trial.

@@

Let $\,b_1, b_2, \ldots \,$ denote a sequence of independent, identically 
distributed, Bernoulli random variables with $\mathbb{P}\{b_i=1\}=p$, for 
probability parameter $0\leq p \leq 1$. In the clinical trial setting 
$\,b_i = 0$ corresponds to an unsuccessful outcome.  Let $s$ and $t$ be 
positive integers.  Define the stopped negative binomial (SNB) random 
variable $Y$ as the smallest 
integer value such that $\,\{b_1, \ldots , b_Y\}\,$ contains {\em either} 
$\,s\,$ successes {\em or} $\,t\,$ failures. The distribution of $\,Y\,$ has 
support on integer values in the range 
\begin{equation*}                                     %   (3)
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}
\end{equation*}
and it is distributed as
\begin{equation} \label{eqn:pmf}
\mathbb{P}[Y=k] = S(k, p, s)  [s \leq k] + T(k, p, t) [t \leq k ] 
\end{equation}
where 
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is the negative binomial probability of $s$ successful outcomes and
\begin{equation} \label{eqn:R}
T(k, p, t) = {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{equation}
is the negative binomial probability of $t$ unsuccessful outcomes. 
From (\ref{eqn:pmf}) one can see that the mass at any point on the support 
of the distribution may come from either $X_k = s$, captured in the 
function $S$, or $X_k = k-t$, captured in the function $T$. The support of 
$S$ is defined by $s \leq S \leq s+t-1$ and the support of $T$ is defined 
by $t \leq T \leq s+t-1$.

%It should be noted that
%this result expressed using the de Finetti where indicator functions and
%sets are equivalent. A more complete explanation of this equivalence can be 
%found in [\ref{Pollard 2002}].

@@

To show the result in Equation (\ref{eqn:pmf}) consider the 
process $\mathbf{X} = \{X(k) : k \in \mathbb{Z}_{\ge 0} \}$ with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + b_{k+1} \left[ k-t < X_k < s \right].
\end{equation*}
As before the $b_{k+1}$'s are distributed as Bernoulli($p$) for all $k$ in the 
domain of $\mathbf{X}$. The process can be conceptualized as a series of coin 
flips that stops when either $s$ heads or $t$ tails are reached. At each step 
a coin is flipped. If it is heads, the process advances one diagonally in the 
positive horizontal and vertical direction. Otherwise, it advances in the 
positive horizontal direction only. When the process stops, $X_k$'s and 
remains either $X_k = s$ or $X_k = k-t$. One example is given in 
Fig. \ref{fig:kane_viz}.

\begin{prop}
The distribution of the stopping time $\argminl_k [X_k \geq s, X_k \leq k-t]$ 
is equal to the SNB and is given in (\ref{eqn:pmf}). 
\end{prop}
\begin{proof}
%The proof will proceed in two parts. First, a combinatorial justification 
%will be given for the probability mass value on each element of the support. 
%Second, it will be shown that the sum of the masses over the support sums to 
%one.

The probability that a given realization of $\mathbf{X}$ reaches $s$ at 
the $k$th outcome is the probability that, at time $k-1$ there are $s-1$ 
successful outcomes and $k-s$ unsuccessful outcomes multiplied by 
the probability of a 
success at time $k$ (\ref{eqn:N}). The probability $X_k = k-t$ 
is the probability that, at outcome $k-1$ there are $k-t$ successful outcomes 
and $t-1$ unsuccessful outcomes multiplied by the probability of an 
unsuccessful outcome at time $k$ (\ref{eqn:R}).  Finally, we need to show that 
the sum
\begin{equation} \label{eqn:sum_proof}
R = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{equation}
is equal to one.
First, substitute $i=k-s$ in the first summation and
$j=k-t$ in the second. 
\begin{equation*}
{j+s-1 \choose s-1} = {j+s-1 \choose j}
\end{equation*}
Then %(\ref{eqn:sum_proof}) becomes:
%\begin{equation*}
%R = \sum_{i=0}^{t-1} {i+s-1 \choose s-1} p^s (1-p)^i +
%\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
%\end{equation*}
$R$ can be written
as the cumulative distribution function of two 
negative binomial distributions
\begin{equation} \label{eqn:transformed_sum}
R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i \; + \;
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation}

Let $I_p(s, t)$ be the {\em regularized incomplete beta function}, which is 
also one minus the c.d.f. of the negative binomial distribution. This
function satisfies $I_p(s, t) = 1-I_{1-p}(t, s)$.  Then it can be proven
that \ref{eqn:pmf} is a valid probability mass by showing that the
probability mass function sums to one.
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-I_p(s, t) + 1 - I_{1-p}(t, s) \\
   &= 1. 
\end{align*}
\end{proof}

\begin{prop} Let $S$ be distributed SNB with parameters $p$, $s$, and $t$.
Then the moment generating function (MGF) of $S$ is
\begin{equation} \label{eqn:mgf}
\mathbb{E}~e^{xS} = \frac{p e^x}{1 - qe^x} I_{qe^x} (s, t) + \frac{qe^x}{1-pe^x} I_{pe^x}(s, t)
\end{equation}
for $q = 1-p$ when $x \leq -\log(q)$ and $x \leq -\log(p)$.
\end{prop}
\begin{proof}
By definition, the MFG of the SNB is:
\begin{equation*}
\mathbb{E}~e^{xS} = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s q^{k-s} e^{kx} 
  + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} q^t e^{kx}
\end{equation*}
and this can be rewritten as:
\begin{equation} \label{eqn:first_sum}
\mathbb{E}~e^{xS} = \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} 
  + \sum_{k=t}^{s+t-1}{k-1 \choose k-t} (qe^x)^t (pe^x)^{k-t}.
\end{equation}
Taking the first summation in Equation \ref{eqn:first_sum}:
\begin{align*}
\sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} &= 
  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose s-1} 
    (qe^x)^{k-s} (1-qe^x)^s) \\
  &= \left(\frac{pe^x}{1 - qe^x}\right)^s I_{qe^x}(s, t).
\end{align*}
Since the incomplete beta function has support on zero to one 
$qe^x \leq 1$. This implies that $x \leq -\log(q)$.

A similar expression can be derived using the same calculation with 
the constraint that $x \leq -\log(p)$. The result
follows from the afore mentioned property of the regularized incomplete
beta function.
\end{proof}

@@

The probability mass function of $\,Y\,$ has a variety of shapes for different choices of the parameters $(s,\, t,\,p)$.
These shapes are illustrated in Fig.~\ref{shapes.fig}.
The SNB is related to the negative binomial distribution. Specifically, if 
$t$ is large then the $Y-s$ has a negative binomial distribution 
with
\begin{equation*}                                    %   (1)
   \mathbb{P}\{Y=s+j \}        \label{nb1.eq}
          = {{s+j-1}\choose{s-1}} p^s (1-p)^j
\end{equation*}
for $\,j=0, 1,\ldots\,$. A similar statement can be made when $s$ is large
and $t$ is small.

For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the 
riff-shuffle, or minimum negative binomial distribution~[\ref{Uppuluri 1970}].
Similar derivations of the closely-related maximum negative binomial discrete 
distributions also appear in~[\ref{Zhang 2000}, \ref{Zelterman 2004}]. 
The maximum negative binomial is the smallest number of outcomes necessary
to observe at least $c$ successes {\em and} $c$ failures, but the SNB give
the number of coin flips to observe {\em either} $s$ heads or $t$ tails.

@@

\section{A Bayesian Analysis}

@@

Let us assume that $p$ has a Beta distribution, with constant prior 
parameters $\alpha$ and $\beta$. Then the closed form 
posterior distribution of the SNB is
\begin{align}
\mathbb{P} \left[Y = k | \mathbf{b}, \alpha, \beta \right] &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
  [s \leq k \leq s+k-1] + \nonumber \\
  & {k-1 \choose k-t} 
  \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
  [t \leq k \leq s+k-1]
\end{align}
where $B$ is the Beta function and $\mathbf{b}$ is the sequence of Bernoulli data.

%When $p$ is known, the probability mass function (PMF) of the number of steps 
%until the process is stopped is given by (\ref{eqn:pmf}). However, many times 
%this is not the case. 
%To do this, we will assume that $p$'s probability in the absence of 
%available data is $\text{Beta}(1/2, 1/2)$, the Jeffrey's prior. This 
%distribution was chosen for two reasons. First, it is the conjugate prior 
%of the (Negative) Binomial distribution and updating the distribution as data 
%are received is straightforward. Second, it is objective and limits the number 
%of ``prior coin-flips'' to one. This means that data are more heavily 
%influence the distributional characteristics of $p$ with more data providing 
%more proportional influence.

\begin{prop}
The posterior PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
\begin{align} \label{eqn:posterior}
f(k | s, t, \alpha, \beta) &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)}     [s \leq k \leq s+k-1] + \nonumber \\
  & {k-1 \choose k-t} 
    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
    [t \leq k \leq s+k-1]
\end{align}
where $B$ is the Beta function.
\end{prop}
\begin{proof}
For notational simplicity, assume that $s,t \leq k \leq s+t-1$. When this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp
\end{align*}
The result follows by applying the definition of the Beta function to the integral terms.
\end{proof}

Equation \ref{eqn:posterior} suggests an alternative construction for 
Phase II clinical trials.  Patients could be 
enrolled sequentially until there are enough patients to guarantee the 
desired level of power or until a specified number of adverse events are 
recorded. In cases such as these the endpoints $s$ and $t$ are known and 
new outcomes can be incorporated into the estimate of
$p$ to provide updated odds of successful enrollment. Later it will be shown 
that estimates of the distribution of $p$ can even be used after an endpoint 
has been reached to determine the amount of uncertainty in the endpoint.

\section{A Bayesian Design for Clinical Trials}

%Returning to the breast cancer trial introduced in the first section, consider
%the first stage, having a sample size of 12. A
%response rate of 2/12 stage corresponds to an individual 
%probability of responding of $p=0.1667$. The standard deviation of this 
%probability is $\sqrt{p(1-p)/12}=0.1076$. If we were to use the Central 
%Limit Theorem then the 95\% confidence interval around the mean would 
%be $[-0.0442, 0.3776]$. While this approach is problematic since the number 
%of samples (12 in this case) is small it does indicate how much uncertainty 
%is built into the first stage. If we consider samples from both stages then 
%the standard deviation drops to 0.0630 and the 95\% confidence interval 
%would be $[0.0432, 0.2901]$.

The Simon Two-stage Optimal Design faces two problems. 
First, success in the first stage is not a good indicator of success 
in the second stage and at the same time failure in first stage in no way 
implies failure in the second stage, especially when $p$ is close to it's 
threshold value. The variance in the estimate is so large that it is a poor 
estimator. The second difficulty is that the design is not {\em enrollment 
efficient}. That is, if the first stage is successful then additional patients 
are enrolled. Enrollment, especially for certain types of cancer, can be very 
difficult and can take time to find eligible patients. Furthermore, if 
all 12 of the patients in the first stage respond positively, should we 
really need to enroll additional patients?

@@

%\item[\bibref{Pollard 2002}]
%Pollard D. {\it A User's Guide to Measure Theoretic Probability}, 
%Cambridge University Press, 2002.
